---
title: "Links inside documents"
vignette: >
  %\VignetteIndexEntry{Links inside documents}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
format:
  html:
    embed-resources: true
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{=html}
<style>
.ll-fig{display:block;margin:1.5rem auto;max-width:100%;height:auto}
.ll-dark{display:none}
@media (prefers-color-scheme:dark){.ll-light{display:none}.ll-dark{display:block}}
[data-bs-theme=light] .ll-light{display:block}
[data-bs-theme=light] .ll-dark{display:none}
[data-bs-theme=dark] .ll-light{display:none}
[data-bs-theme=dark] .ll-dark{display:block}
</style>
```

```{r setup}
library(livelink)
```

If you write in Quarto or R Markdown, you should not have to build a link in one chunk
and paste the URL into another. livelink plugs into knitr, so a chunk can carry its own
link.

There are two ways in, and the choice comes down to one question: **should the code
also run in your document?**

![The hook leaves the chunk running, so the reader sees the output and gets a link. The engine replaces execution: the code is shown, nothing runs, and only a link is produced.](../man/figures/chunk-modes-light.svg){.ll-fig .ll-light}
![](../man/figures/chunk-modes-dark.svg){.ll-fig .ll-dark aria-hidden="true"}

# The hook: output *and* a link

Add `livelink: true` to an ordinary `r` chunk. Nothing else changes. The chunk runs,
its output appears, and a link is added underneath.

````
```{{r}}
#| livelink: true
#| autorun: true
# Load the data
data(mtcars)
plot(mtcars$mpg, mtcars$wt)   # a comment your reader will actually see
```
````

That is a real chunk. Here it is, running in this very vignette. The plot below is
drawn by R, here, and the link under it opens the same code in the browser:

```{r}
#| livelink: true
#| autorun: true
#| panels: ["editor", "plot"]
# Load the data
data(mtcars)
plot(mtcars$mpg, mtcars$wt)   # a comment your reader will actually see
```

This is usually what you want. Readers see the result, and can still go and poke at it.

# The engine: a link, and no execution

Sometimes the code must *not* run in your document: a Shiny app, something needing a
package you have not installed, something slow. Use a ```` ```{livelink} ```` chunk.
The code is displayed, a link is produced, and nothing is executed.

````
```{{livelink}}
#| engine.target: shinylive-r
#| mode: app
library(shiny)
shinyApp(fluidPage(sliderInput("n", "n", 1, 100, 50)), function(input, output) {})
```
````

That is how the Shiny examples in these vignettes are built. livelink does not depend
on shiny, and does not want to.

::: {.callout-note}
## There is no `{shinylive-r}` chunk

It is tempting to want a ```` ```{shinylive-r} ```` chunk to match Shinylive's own
naming. knitr cannot provide one: its chunk syntax forbids a hyphen in an engine name,
and in Quarto a `{shinylive-r}` cell is handed to the Shinylive extension, not to
knitr. Name Shinylive through `engine.target` on a `{livelink}` chunk instead, as
above.
:::

# Chunk options

Each option changes one part of the link. The figure below reads left to right: the
option you set, the piece of the URL it changes, and what that means for your reader.
The full list is at the end of this article, in [Every chunk
option](#every-chunk-option).

![Each option changes one part of the URL. `echo` is the exception: it changes what the reader sees, not the link.](../man/figures/chunk-options-light.svg){.ll-fig .ll-light}
![](../man/figures/chunk-options-dark.svg){.ll-fig .ll-dark aria-hidden="true"}

## `echo` does not do what you might think

It is natural to assume the code must be visible for a link to be made. It does not.

`echo` controls whether the **source is displayed in your page**. The link is built
from the chunk's source, which knitr hands over regardless. So `echo: false` gives you
a working link whose code you simply cannot see in the document:

```{r}
#| livelink: true
#| echo: false
#| link.text: "This link works, though you cannot see its code above"
mean(mtcars$mpg)   # this comment still travels in the link
```

The code, and that comment, are both in the link. Follow it and look.

`eval: false` is the other half: the chunk is shown but not run, which makes an `r`
chunk behave rather like the engine.

# Setting options once

Chunk options are knitr options, so `opts_chunk` sets them for the whole document:

```{r}
#| eval: false
knitr::opts_chunk$set(livelink = TRUE, autorun = TRUE)
```

Every subsequent `r` chunk then carries a link, and you opt one out where you do not
want it:

````
```{{r}}
#| livelink: false
# scaffolding the reader does not need a link for
setup_the_data()
```
````

This is worth doing in course notes, where nearly every chunk should be something a
student can open and edit.

# Why not a braced expression?

`webr_repl_link({ ... })` looks like it ought to work in a document. It does, but it
loses your comments.

knitr runs chunks through `evaluate::evaluate()`, which discards **source
references**, and comments live in those. So a `{ }` expression inside a knitted
document arrives at livelink with the comments already gone, and no `keep.source`
setting brings them back.

Both the hook and the engine are handed the chunk's **raw source**, which is the only
route by which a comment reaches a link from a rendered document. That is the whole
reason they exist.

Use braces in an interactive session, where source references are present. Inside a
document, use a chunk.

# Every chunk option {#every-chunk-option}

The same shape as the figure above, in full. Each option, the part of the link it
changes, and what that means for your reader.

| Option | In the link | What it means |
|---|---|---|
| `livelink: true` | `webr.r-wasm.org/…` | hook only: turn an `r` chunk into a webR link |
| `livelink: shinylive-r` | `shinylive.io/r/…` | hook only: a Shiny app instead of a REPL (also `"webr"`, `"shinylive-py"`) |
| `engine.target: shinylive-r` | `shinylive.io/r/…` | engine only: the same targets, for a `{livelink}` chunk |
| `autorun: true` | `&jza` (the `a` flag) | webR only: run the code the moment the link opens |
| `panels: ["editor", "plot"]` | `?mode='editor-plot'` | webR only: which panels appear |
| `mode: app` | `shinylive.io/r/app/` | Shinylive only: the running app, not the code |
| `filename: analysis.R` | `analysis.R` | the file's name inside the environment |
| `link.text: "Try it"` | `[Try it](…)` | the words the reader clicks |
| `link.only: true` | *(unchanged)* | engine only: emit the link without showing the source |
| `echo: false` | *(unchanged)* | hides the code in the page; the link still works |

Any of these can be set once for a whole document, as [Setting options
once](#setting-options-once) shows.
