---
title: "Multi-file projects and Shiny apps"
vignette: >
  %\VignetteIndexEntry{Multi-file projects and Shiny apps}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
format:
  html:
    embed-resources: true
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r setup}
library(livelink)
```

Real work is rarely one file. This article covers the multi-file half of livelink:
webR projects, Shinylive apps in R and Python, and batch-processing a directory of
them.

## Two ways to describe a project

Every multi-file function accepts the same two input shapes, and you can use whichever
is closer to hand.

A **named list**, when the code is already in R:

```{r}
project <- list(
  "main.R"    = "source('utils.R')\nsummarise(mtcars)",
  "utils.R"   = "summarise <- function(d) summary(d)",
  "README.md" = "# Analysis\n\nRun main.R to start."
)

webr_repl_project(project)
```

The list names become the filenames inside the environment. Non-R files are welcome --
data, documentation, anything the code needs.

A **vector of file paths**, when the code is on disk:

```{r}
dir <- file.path(tempdir(), "project")
dir.create(dir, showWarnings = FALSE)
writeLines("source('utils.R')", file.path(dir, "main.R"))
writeLines("f <- function() 42", file.path(dir, "utils.R"))

webr_repl_project(c(file.path(dir, "main.R"), file.path(dir, "utils.R")))
```

Files are read and named by their basename, so the paths on your machine do not leak
into the link.

## Running something on arrival

`autorun_files` names the files to execute when the link opens.

```{r}
link <- webr_repl_project(project, autorun_files = "main.R")
link$url |> substr(1, 60)
```

Pass `"all"` to run every R file in the project:

```{r}
webr_repl_project(project, autorun_files = "all")
```

Naming a file that is not in the project is an error rather than a silent no-op:

```{r}
#| error: true
webr_repl_project(project, autorun_files = "nonexistent.R")
```

## Shiny apps

`shinylive_r_link()` takes an app the same three ways: a string, a file path, or a
named list of files.

A single-file app is the common case, and a bare string becomes `app.R`:

```{r}
app <- "
library(shiny)

ui <- fluidPage(
  titlePanel('Old Faithful'),
  sliderInput('bins', 'Bins:', 1, 50, 30),
  plotOutput('hist')
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(faithful$waiting, breaks = input$bins, col = 'steelblue')
  })
}

shinyApp(ui, server)
"

shinylive_r_link(app)
```

For an app split across files, pass a named list:

```{r}
shinylive_r_link(list(
  "app.R"   = "library(shiny)\nsource('ui.R')\nsource('server.R')\nshinyApp(ui, server)",
  "ui.R"    = "ui <- fluidPage(titlePanel('Split app'))",
  "server.R" = "server <- function(input, output) {}"
))
```

### Editor or app?

`mode` decides what a visitor lands on.

```{r}
# "editor" -- code and app side by side. Good for teaching and code review.
shinylive_r_link(app, mode = "editor")
```

```{r}
# "app" -- the running application only. Good for end users.
shinylive_r_link(app, mode = "app")
```

In `app` mode you can also drop the Shinylive header, which is what you want when the
app is going into an `<iframe>` on someone else's page:

```{r}
shinylive_r_link(app, mode = "app", header = FALSE)
```

Note that `mode` here means something different from webR's `panels`. Shinylive has
two display modes; webR has four panels you can show or hide. They are separate
arguments because they are separate ideas.

### Python

Python Shiny apps work identically, and a bare string becomes `app.py`:

```{r}
py_app <- "
from shiny import App, render, ui

app_ui = ui.page_fluid(
    ui.h2('Hello from Python'),
    ui.output_text('greeting'),
)

def server(input, output, session):
    @render.text
    def greeting():
        return 'Running in the browser, no server required.'

app = App(app_ui, server)
"

shinylive_py_link(py_app, mode = "app")
```

`shinylive_project()` is the explicit form when you want to name the engine yourself:

```{r}
shinylive_project(
  list("app.py" = py_app, "utils.py" = "def helper():\n    return 42"),
  engine = "python"
)
```

## A directory of apps

If you keep each app in its own folder -- the usual Shiny layout -- `shinylive_directory()`
turns the lot into links in one pass.

```{r}
apps <- file.path(tempdir(), "apps")
dir.create(file.path(apps, "histogram"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(apps, "scatter"),   recursive = TRUE, showWarnings = FALSE)
writeLines(app, file.path(apps, "histogram", "app.R"))
writeLines(app, file.path(apps, "scatter", "app.R"))

links <- shinylive_directory(apps, engine = "r", mode = "app")
links
```

Each subdirectory containing an `app.R` (or `app.py`, for Python) becomes one link,
named after the folder. Every other file in that folder -- helpers, data, CSS -- is
bundled into the app.

```{r}
repl_urls(links)
```

A directory with no apps in it warns and hands back an empty result, rather than
failing:

```{r}
empty <- file.path(tempdir(), "no-apps")
dir.create(empty, showWarnings = FALSE)

shinylive_directory(empty, engine = "r")
```

## How big can a link get?

The files travel *inside* the URL, so a project's size is bounded by what a browser
will accept in a URL -- in practice tens of kilobytes of code, which is a lot of R.

The payload is gzipped before it is encoded, so ordinary source code compresses well.
Large binary data is the wrong thing to put in a link; point at it from the code
instead.
